---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, cache = TRUE)
```

## Loading and preprocessing the data

Download dataset if necessary.
```{r}
data_url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if (!file.exists("activity.csv")) {
        if (file.exists("activity.zip")) {
                download.file(url = data_url, destfile = "activity.zip", method = "curl")
        }
        unzip("activity.zip")
}
```
Load needed package and dataset.
```{r, eval=FALSE}
library(tidyverse)
```

```{r}
activity <- read_csv("activity.csv")
activity
```


## What is mean total number of steps taken per day?

### Histogram of the total number of steps taken each day
```{r}
daily_steps <- activity %>%
        group_by(date) %>%
        summarise(total_steps = sum(steps))
daily_steps

ggplot(daily_steps, aes(total_steps)) +
        geom_histogram() + 
        labs(title = "Daily Steps", x = "Steps", y = "Frequency")
```

### Mean and median number of steps taken each day
```{r}
mean(daily_steps$total_steps, na.rm = TRUE)
median(daily_steps$total_steps, na.rm = TRUE)
```


## What is the average daily activity pattern?
```{r}
average_interval_steps <- activity %>%
        group_by(interval) %>%
        summarise(average_steps = mean(steps, na.rm = TRUE))

ggplot(average_interval_steps, aes(interval, average_steps)) + 
        geom_line() +
        labs(title = "Daily Acitivity Pattern", x = "Interval", y = "Average Steps")
        
```

Find the interval which contain max average steps.
```{r}
average_interval_steps[which.max(average_interval_steps$average_steps),]
```


## Imputing missing values

### Calculate the total number of missing values in the dataset 
```{r}
table(is.na(activity$steps))
```

### Strategy for filling in all of the missing values in the dataset
According to `daily_steps`, step data are missed in some whole days. We can simply fill the missing data in the with steps of `average_interval_steps`.


### Create a new dataset with missing data filled
```{r}
na_date <- filter(daily_steps, is.na(total_steps))[["date"]]
activity[is.na(activity$steps), "steps"] <- rep(average_interval_steps$average_steps, times = length(na_date))
```
Let's check NAs again
```{r}
table(is.na(activity$steps))
```


### Histogram of the total number of steps taken each day
```{r}
daily_steps <- activity %>%
        group_by(date) %>%
        summarise(total_steps = sum(steps))
daily_steps

ggplot(daily_steps, aes(total_steps)) +
        geom_histogram() + 
        labs(title = "Daily Steps", x = "Steps", y = "Frequency")
```


### Calculate the mean and median total number of steps taken per day
```{r}
mean(daily_steps$total_steps)
median(daily_steps$total_steps)
```


## Are there differences in activity patterns between weekdays and weekends?
Create a new variable indicating each observation is a weekday or weekend and plot activity pattern.
```{r}
activity <- activity %>% 
        mutate(day_of_week = weekdays(date))

activity$day_of_week <- sub("Monday|Tuesday|Wednesday|Thursday|Friday", "weekday", activity$day_of_week)
activity$day_of_week <- sub("Saturday|Sunday", "weekend", activity$day_of_week)

```

```{r}
average_interval_steps <- activity %>%
        group_by(interval, day_of_week) %>%
        summarise(average_steps = mean(steps))
average_interval_steps

ggplot(average_interval_steps, aes(interval, average_steps, color = day_of_week)) + 
        geom_line() +
        facet_wrap(.~ day_of_week, nrow = 2) +
        labs(title = "Daily Acitivity Pattern", x = "Interval", y = "Average Steps") +
        theme(legend.title=element_blank())

```

